---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# indexvctrs

The goal of indexvctrs is to provide a DSL for indexed vector operations

Just experimental -- I don't expect this to be usable any time soon.

## Installation

``` r
devtools::install_github("jameelalsalam/indexvctrs")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example, error=TRUE}
with_indices({
  Emissions[ghg, fuel] <- fuel_consumption[fuel] * EF[ghg, fuel]
  
  Emissions[ghg] <- sum_over(fuel, Emissions[ghg, fuel])
})
```

Example equation 2.1 from [IPCC 2006 inventory GLs](https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/2_Volume2/V2_2_Ch2_Stationary_Combustion.pdf)

Under the hood, these objects are actually data frames with additional attributes indicating which column is the `value` column and which columns are the indices. Unlike 'normal' vector math, which does matching and recycling by position, the DSL operations match and recycle along corresponding indices, akin to `dplyr::full_join()`.

I don't believe that R really does what I want when recycling vectors and arrays or arrays of different dimension.

```{r, error=TRUE}
c(1:2) + array(c(11:16), dim = c(2, 3))
array(c(1:2), dim = 2) + array(c(11:16), dim = c(2, 3))
```


Representation invariant:

* To simulate an nd-array, the data must be unique on the combination of indices.
* Only a single 'value' column is allowed

## References

I hope to take advantage of learning from:

* [DSLs](https://adv-r.hadley.nz/translation.html) in adv-R
* S3 object creation from [adv-R](https://adv-r.hadley.nz/s3.html), the [vctrs pacakge](https://github.com/r-lib/vctrs), and the [sloop package](https://github.com/hadley/sloop)
* related to Einstein summation notation, e.g. in [Juilia](https://github.com/ahwillia/Einsum.jl) or [Python/Numpy](https://docs.scipy.org/doc/numpy-1.10.0/reference/generated/numpy.einsum.html)


