# ---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# indexvctrs

<!-- badges: start -->
[![R build status](https://github.com/jameelalsalam/indexvctrs/workflows/R-CMD-check/badge.svg)](https://github.com/jameelalsalam/indexvctrs/actions)
[![R-CMD-check](https://github.com/jameelalsalam/indexvctrs/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jameelalsalam/indexvctrs/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of {indexvctrs} is to provide a DSL for indexed vector operations

Pretty experimental right now!

## Installation

``` r
devtools::install_github("jameelalsalam/indexvctrs")
```

## Example

This package provides `idx_tbl` objects which behave somewhat like sparse multi-dimensional arrays with named indices, but remain data frames as well. One of the columns is the `value` column, which is manipulated directly through math operations. 

When `idx_tbl` objects are created, columns are marked as index columns (the idx_cols attribute), and a single value column is stored as `value`. Additional columns are dropped.

```{r}
library(indexvctrs)

crop_acres <- idx_tibble(
  tibble::tribble(
    ~crop,   ~year, ~value,
    "corn",  2005,  4,
    "wheat", 2005,  5,
    "corn",  2010,  6,
    "wheat", 2010,  8
  ),

  idx_cols = c("crop", "year")
)

EF <- idx_tibble(
  tibble::tribble(
    ~crop, ~value,
    "corn", 1.5,
    "wheat", 2
  ), idx_cols = "crop"
)

idx_cols(crop_acres)
crop_acres
```

Math operations can be performed between idx_tbls with common indices via join semantics. The result of this is that where indices do not share an axis, values are broadcast (or recycled) across that axis. This applies, for example, to scalars.

```{r}
EF * crop_acres
```

```{r}
EF * 2
```


It can be convenient to express relationships using functions (e.g., so that they can be stated out-of-order):

```{r}
calc_emissions <- function(activity, EF) {activity * EF}

emissions_by_crop <- calc_emissions(activity = crop_acres, EF = EF * 2)
emissions_by_crop
```

Vector operations act on the `value` column:

```{r}
sum(emissions_by_crop)
```


If something isn't implemented, never fear because you should be able to drop back to data frame operations.

```{r}
library(tidyverse)

emissions_by_crop %>%
  filter(year == 2010)

```


## References

I hope that this package benefits from learning from some of these:

* [DSLs](https://adv-r.hadley.nz/translation.html) in adv-R
* S3 object creation from [adv-R](https://adv-r.hadley.nz/s3.html), the [vctrs pacakge](https://github.com/r-lib/vctrs), and the [sloop package](https://github.com/hadley/sloop)
* exploration of recycling rules in the [{rray} package](https://rray.r-lib.org/articles/broadcasting.html)
* related to Einstein summation notation, e.g. in [Juilia](https://github.com/ahwillia/Einsum.jl) or [Python/Numpy](https://docs.scipy.org/doc/numpy-1.10.0/reference/generated/numpy.einsum.html)


